---
title: "Report of CCVR04"
author: "Tom Beesley"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE, message = FALSE)
date_compiled_title = c("Output created on", date())
```
---
date: `r date_compiled_title`
---

```{r}
library(tidyverse)
library(broom)
library(patchwork)
library(knitr)
library(kableExtra)
library(afex)
library(BayesFactor)

theme_set(theme_classic())

setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # set wd to current file location

data <- read_csv("data_tidy_CCVR04.csv") # read in the tidied data from the first analysis script.

data_AW <- 
  data %>% 
  filter(phase == 2) %>% # separate out awareness data
  mutate(TT = if_else(set %in% c(5,6), "Repeated", "Random"), # recode "set" variable into 2 variables
         targ_depth = if_else(set %in% c(5,7), "Near targets", "Far targets")) %>% 
  select(exp:block, set, TT, targ_depth, everything())

dataMain <- 
  data %>% 
  filter(phase == 1) %>% # main phase of the task
  mutate(TT = if_else(set %in% c(1,2), "Repeated", "Random"), # recode "set" variable into 2 variables
         targ_depth = if_else(set %in% c(1,3), "Near", "Far"),
         epoch = ceiling(block/2)) %>% 
  select(exp:phase, epoch, block:set, TT, targ_depth, everything())

```

This file takes the form of a brief "lab report", but is also the main data processing of CCVR04, which feeds into the CCVR_ms_1 (manuscript) folder

**Summary of the DESIGN of the experiment**  

The main control of the pattern generation is found in "CreatePats.m" in the experiment code folder

Set 1 - Near target repeated configurations (half pattern repeated, half random)  
Set 2 - Far target repeated configurations (half pattern repeated, half random)  
Set 3 - Near target random configurations  
Set 4 - Far target random configurations  

odd participants: proximal distractors were repeated, distal Ds randomised   
even participants: distal distractors were repeated, proximal Ds randomised  

Blocks 1:16, 21:36, 41:56, 61:62 - Repeated configurations (sets 1 and 2)  
Blocks 17:20, 37:40, 57:60 - Random configurations (sets 3 and 4)  

NB: the block numbers in the raw data files were mislabeled - blocks 21:36 were written out as 20:35. This had no consequence on the presentation of stimuli, but the block numbers are adjusted in 1_read_tidy_CCVR04.R  

**Analysis report** 

```{r}
# a table of the demographic information across participants
data_demog <- 
  dataMain %>% 
  group_by(subj) %>% 
  slice(1) %>% # take top row
  ungroup() %>% 
  select(subj,age,gender,condition)

demog_summary <- 
  data_demog %>% 
  summarise(n = n(),
            mean_age = mean(age),
            sd_age = sd(age),
            males = sum(gender=="M"),
            females = sum(gender=="F"),
            non_binary = sum(!gender %in% c("M", "F")))

# summary statistics of response patterns,
basic_resp_stats <- 
  dataMain %>%
  group_by(subj) %>%
  summarise(mean_RT = mean(RT < 9999),
            perc_Timeouts = sum(dist_to_T==-1)/n(),
            mean_Extra_Resp = mean(num_additional_Rs),
            perc_Rs_Over1 = sum(num_additional_Rs>0)/n(),
            perc_Rs_Over2 = sum(num_additional_Rs>1)/n()) %>% 
  summarise_all(list(mean = mean, sd = sd)) %>%  # from participant summaries, compute summary stats
  select(-contains("subj"))

# create table from those summary statistics
table_basic_resp_stats <- 
  basic_resp_stats %>% 
  pivot_longer(cols = everything(), names_to = "stat") %>% 
  kable(digits = 2, col.names = c("","")) %>% 
  kable_styling(latex_options = "hold_position", position = "left")

# filter trials with more than 2 responses (num additional Rs > 1), timeouts,
# and outlier RTs  
data <- 
  dataMain %>% 
  filter(num_additional_Rs<=1, dist_to_T>=0) %>% 
  group_by(subj) %>% 
  mutate(zRT = scale(RT)) %>% 
  filter(between(zRT,-2.5,2.5)) %>% 
  ungroup() 

# what percentage of trials were removed for each participant?
perc_removed <- 
  data %>%
  group_by(subj) %>% 
  summarise(perc = (1 - (n()/496))*100)

# compute normalised RT (norm_RT), for calculating w/s error bars 
data <- 
  data %>% 
  group_by(subj) %>% # within each experiment, and each subject
  summarise(sub_MeanRT = mean(RT)) %>% # calculate the mean RT for the subject
  mutate(sample_MeanRT = mean(sub_MeanRT)) %>% # the overall mean RT for the sample in that experiment
  ungroup() %>% # ungroup the data, to allow for removal of variable
  left_join(data, by = "subj") %>% # join to the original data frame
  mutate(norm_RT = RT - sub_MeanRT + sample_MeanRT) %>% # compute the normalised RT, from the RT, subject mean and sample mean
  select(exp, subj, age:RT, norm_RT, everything(), -sub_MeanRT, -sample_MeanRT) # arrange the variables

```

There were `r length(data_demog$subj)` participants in the sample, of which `r sum(data_demog$gender == "F")` identified as female, and `r sum(data_demog$gender == "M")` as male. `r sum(data_demog$condition == "proximal_repeated")` were assigned to the "proximal repeated" condition, while `r sum(data_demog$condition == "distal_repeated")` were assigned to the "distal repeated" condition.

Summary statistics for RTs and number of responses made:

`r table_basic_resp_stats`

Data processing: trials which led to timeouts, and trials that had more than one additional responses (i.e., up to 2 responses allowed) were removed. Following this, RTs greater or less than 2.5 SDs from the participant mean RT were removed. On average this resulted in the loss of `r round(mean(perc_removed$perc),1)`% of trials. Normalised RT was computed in order to create within-subject error bars in all plots. 

```{r}
# summarise the overall mean RT for each P 
overall_RT_Means <- 
  data %>% 
  group_by(subj) %>% 
  summarise(meanRT = mean(RT)) %>% 
  mutate(z_meanRT = scale(meanRT))

outlier_meanRTs <- 
  overall_RT_Means %>% 
  filter(!between(z_meanRT, -2.5, 2.5)) %>% 
  pull(subj) # Ps 27 and 60 seems to have unusually long mean RTs
  
# plot the boxplots of RT
RT_boxplot <- 
  overall_RT_Means %>% 
  ggplot(aes(x = "meanRT", y = meanRT))+
  geom_boxplot() +
  geom_point() +
  geom_text(aes(label = subj), hjust = 2, check_overlap = TRUE)

# remove the two outlier participants from the sample
data <- filter(data,!subj %in% outlier_meanRTs)

```
Mean RTs for each participant were computed and the mean across the sample was `r round(mean(overall_RT_Means$meanRT),0)` ms (SD = `r round(sd(overall_RT_Means$meanRT),0)`). `r length(outlier_meanRTs)` participants [`r outlier_meanRTs`] produced means that were more than 2.5 SDs above the mean of the sample, and were removed from further analysis. 


```{r}
# exploring the difficulty of target detection at different depths

# how many trials contributed to the analysis by target depth

data %>% 
  group_by(subj, targ_depth) %>% 
  summarise(prop = n()/240) %>% 
  group_by(targ_depth) %>% 
  summarise(meanProp = mean(prop))

# any unusual participants?

depth_difficulty_measure <- 
  data %>% 
  group_by(subj, targ_depth) %>% 
  summarise(prop = n()/240) %>%
  pivot_wider(names_from = targ_depth, values_from = prop) %>% 
  mutate(discrepancy = Near-Far) 

depth_difficulty_measure %>% 
  ggplot(aes(discrepancy)) +
  geom_density()

# remove those with large discrepancy
depth_difficulty_outliers <- 
  depth_difficulty_measure %>% 
  filter(!between(discrepancy,-.15,.15)) %>% 
  pull(subj)

# remove the depth difficulty outliers from the sample
data <- filter(data,!subj %in% depth_difficulty_outliers)

# remove them from the measure of depth difficulty
depth_difficulty_measure <- 
  depth_difficulty_measure %>% 
  filter(between(discrepancy,-.15,.15)) 

```



RTs were analysed by averaging the data across two consecutive blocks, producing 31 epochs of XX trials. As can be seen from the figure, RTs were slower for near targets compared to far targets. At three points in the experiment, random trials were introduced (epochs 9 and 10, 19 and 20, 29 and 30). These seemed to cause the expected increase in RT, which is evident particularly later on in the experiment.

```{r include = TRUE, warning=FALSE}
# main analysis of RT

# convert TT and targ_depth to factors and set level order
data <- 
  data %>%
  mutate(TT = factor(TT, levels = c("Repeated", "Random")),
         targ_depth = factor(targ_depth, levels = c("Near", "Far")))

legendLBLs <- c("Repeated Near T", "Random Near T", "Repeated Far T", "Random Far T")

# get data for the RT figure. 
data_RT_avgs_figure <- 
  data %>%
  group_by(condition, TT, targ_depth, epoch) %>% 
  summarise(meanRT = mean(RT), 
            SE = sd(norm_RT)/sqrt(n())) %>%
  mutate(condition = recode(condition, 
                            "distal_repeated" = "Distal condition",
                            "proximal_repeated" = "Proximal condition")) %>% 
  ungroup() %>% 
  complete(TT, targ_depth, epoch, condition) # add NA values where missing - ensures breaks in lines for plot

# main RT figure
RT_avgs_figure <- 
  data_RT_avgs_figure %>%
  ggplot(aes(x = epoch, y = meanRT, group = interaction(TT, targ_depth))) +
  geom_line() +
  geom_errorbar(aes(ymin = meanRT-SE, ymax = meanRT+SE), width = 0.2, size = .5) +
  geom_point(aes(fill = interaction(TT, targ_depth),
                 shape = interaction(TT, targ_depth)),
             size = 2) +
  facet_wrap(~condition, dir = "v", scales = "free") +
  theme(strip.text = element_text(size = 12, face = "bold"),
        strip.text.x = element_text(hjust = 0.05),
        strip.background = element_blank(),
        panel.spacing.y = unit(2, "lines")) +
  scale_y_continuous(limits = c(1400, 3000),
                     breaks = seq(1400,3000,200)) +
  scale_x_continuous(breaks = seq(1,31,4),
                     minor_breaks = seq(1,31,2)) +
  scale_fill_manual(name  = "",
                    labels = legendLBLs,
                    values = c("black", "white", "black", "white")) +
  scale_shape_manual(name = "",
                     labels = legendLBLs,
                     values = c(21,21,22,22)) +
  labs(y = "Mean response time (ms)",
       x = "Epoch (16 trials)")

RT_avgs_figure # print the figure
```

The difference in RT between repeated and random patterns during "test phases" (Test 1 = epochs 8:11; Test 2 is epochs 18:21; Test 3 is epoch 28:31)

```{r include = TRUE}
# get the data from the "test" epochs
data_test <- 
  data %>%
  mutate(test = case_when(epoch %in% 8:11 ~ 1,
                          epoch %in% 18:21 ~ 2,
                          epoch %in% 28:31 ~ 3)) %>% 
  filter(test > 0) #remove data not in the test periods
  
# plot the test data by condition
test_figure <- 
  data_test %>% 
  mutate(condition = recode(condition, 
                            "distal_repeated" = "Distal contingencies", 
                            "proximal_repeated" = "Proximal contingencies"),
         targ_depth = recode(targ_depth,
                             "Near" = "Near targets",
                             "Far" = "Far targets")) %>%
  group_by(condition, TT, targ_depth, test) %>%
  summarise(meanRT = mean(RT), 
            SE = sd(norm_RT)/sqrt(n())) %>% 
  ggplot(aes(x = test, y = meanRT, fill = TT)) +
  geom_col(position = position_dodge(),
           colour = "black",
           size = 1) +
  geom_errorbar(aes(ymin = meanRT, ymax = meanRT+SE), 
                width = 0.2, 
                size = .5, 
                position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(breaks = seq(1400,2800,200)) +
  coord_cartesian(ylim = c(1400,2800)) +
  facet_wrap(vars(condition,targ_depth), 
             nrow = 2, 
             scales = "free",
             labeller = labeller(.rows = label_value, .multi_line = TRUE)) + 
  theme(strip.text = element_text(size = 10, face = "bold"),
        strip.text.x = element_text(hjust = 0),
        strip.background = element_blank(), 
        panel.spacing.y = unit(2, "lines")) +
  labs(x = "Test phase",
       y = "Mean RT",
       fill = "")

test_figure #print figure

```

The data from the Test phases were submitted to ANOVA:

```{r include = TRUE}

data_test$subj <- as.factor(data_test$subj)

#get data for the ANOVA on the test data
data_RT_ANOVA <- 
  data_test %>% 
  group_by(condition, subj, TT, targ_depth, test) %>% 
  summarise(meanRT = mean(RT)) 

# conduct ANOVA
data_RT_ANOVA %>% 
  aov_car(meanRT ~ Error(subj/(test*TT*targ_depth)) + condition, data = .)


```

The main effect of condition (proximal or distal distractors repeating) was not significant, and did not interact with any other factor. 

The trial type by target depth interaction was not significant. This suggests that the learning was equivalent across depths. However, the ANOVA reveals an interesting test by trial type by target depth interaction effect, which is significant. This suggests that the pattern of learning effects observed across three tests was quite different in the two depths. Eye-balling the above figure suggests this is because there is a gradual emergence of a CC effect for near targets, but that the effect is present in Test 1 for far targets. Further, the CC effect seems larger in Test 3 for near targets than far targets. 

To explore this interaction further we can remove the TT factor by subtracting RTs for repeated patterns from those for random patterns:

```{r include = TRUE}

# compute a new CC_RT variable, which simplifies the random-repeated comparison
test_diff <- 
  data_test %>%
  group_by(subj, TT, targ_depth, test) %>%
  summarise(meanRT = mean(RT)) %>% 
  pivot_wider(names_from = TT, values_from = meanRT) %>% 
  mutate(CC_RT = Random - Repeated) 

# plot the CC effect as a function of target depth and test.
test_diff_figure <- 
  test_diff %>% 
  group_by(targ_depth, test) %>% 
  summarise(RT = mean(CC_RT), SE = sd(CC_RT)/sqrt(n())) %>% 
  ggplot(aes(x = test, y = RT)) +
  geom_errorbar(aes(ymin = RT-SE, ymax = RT+SE), 
                width = 0.1, size = .5, colour = "black") +
  geom_line(aes(group = targ_depth), size = 1) +
  geom_point(aes(fill = targ_depth), 
             size = 4, shape = 21, colour = "black", stroke = 1.5) +
  scale_x_continuous(breaks = 1:3) +
  scale_y_continuous(limits = c(-100, 300), breaks = seq(-100,300,50)) +
  scale_fill_discrete(type = c("black", "white")) +
  labs(x = "Test",
       y = "Difference in RT (Random - Repeated)",
       fill = "")

# draw figure
test_diff_figure

```

We can look at t-tests on the three tests to see whether these differences were real.

```{r include = TRUE}

# t-test Test 1
test_diff %>% 
  filter(test == 1) %>% 
  {t.test(.$CC_RT ~ .$targ_depth)}

# t-test Test 2
test_diff %>% 
  filter(test == 2) %>% 
  {t.test(.$CC_RT ~ .$targ_depth)}

# t-test Test 3
test_diff %>% 
  filter(test == 3) %>% 
  {t.test(.$CC_RT ~ .$targ_depth)}


```

It does look like there was a real difference in Test 1 and Test 3.

## ego-centric analysis

We have explored the allocentric contingencies in the design, but we can also look at whether the learning varied as a factor of whether the contingencies were close to the observer or not, which we term the egocentric factor.

```{r include = TRUE}

# compute new variables based on egocentricity of the repeated distractors
data_ego_test <- 
  data_test %>%
  mutate(egocentric = case_when(condition == "proximal_repeated" & targ_depth == "Near"
                                ~ "Egocentric - Near",
                                condition == "proximal_repeated" & targ_depth == "Far"
                                ~ "Egocentric - Far",
                                condition == "distal_repeated" & targ_depth == "Near"
                                ~ "Egocentric - Far",
                                condition == "distal_repeated" & targ_depth == "Far"
                                ~ "Egocentric - Near")) %>% 
  mutate(egocentric = fct_relevel(egocentric, "Egocentric - Near", "Egocentric - Far"))

# plot mean RTs as a function of egocentricity, trial type, and test.
test_ego_figure <- 
  data_ego_test %>% 
  group_by(egocentric, TT, test) %>%
  summarise(meanRT = mean(RT), SE = sd(norm_RT)/sqrt(n())) %>% 
  ggplot(aes(x = test, y = meanRT, fill = TT)) +
  geom_col(position = position_dodge(),
           colour = "black",
           size = 1) +
  geom_errorbar(aes(ymin = meanRT, ymax = meanRT+SE), 
                width = 0.2, size = .5, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(breaks = seq(1800,2400,100)) +
  coord_cartesian(ylim = c(1800,2400)) +
  facet_wrap(vars(egocentric), 
             nrow = 2, 
             scales = "free",
             labeller = labeller(.rows = label_value, .multi_line = FALSE)) + 
  theme(strip.text = element_text(size = 10, face = "bold"),
        strip.text.x = element_text(hjust = 0.05),
        strip.background = element_blank(), 
        panel.spacing.y = unit(2, "lines")) +
  labs(x = "Test phase",
       y = "Mean RT",
       fill = "")

test_ego_figure

```

A repeated-model ANOVA shows that the factor of egocentricity of the configuration (near to observer or far from the observer) has no bearing on response times or the TT effect):

```{r include = TRUE}

# convert variables to factors
data_ego_test$subj <- as.factor(data_ego_test$subj)
data_ego_test$egocentric <- as.factor(data_ego_test$egocentric)

# summarise data for ANOVA
data_ego_RT_ANOVA <- 
  data_ego_test %>% 
  group_by(subj, egocentric, TT, test) %>% 
  summarise(meanRT = mean(RT)) 

# conduct ANOVA
data_ego_RT_ANOVA %>% 
  aov_car(meanRT ~ Error(subj/(egocentric*TT*test)), data = .)



```

```{r}
# awareness data analysis - this is not reported here, but is reported in the main manuscript

# remove the outlier participants from the sample
data_AW <- 
  data_AW %>% 
  filter(!subj %in% outlier_meanRTs,
         !subj %in% depth_difficulty_outliers)

# summarise awareness test accuracy as a function of trial type and target depth
data_gen_descriptives <- 
  data_AW %>% 
  group_by(TT, targ_depth) %>% 
  summarise(acc = mean(dist_to_T),
            acc_sd = sd(dist_to_T))

# plot those means
data_gen_descriptives %>% 
  ggplot(aes(x = targ_depth, y = acc, fill = TT)) +
  geom_col(position = position_dodge2(), colour = "black") +
  scale_fill_manual(values = c("black","white"))

# subject level means of awareness accuracy, as a function of trial type and target depth 
data_gen_analysis <- 
  data_AW %>% 
  group_by(subj, TT, targ_depth) %>% 
  summarise(acc = mean(dist_to_T))

# look at the correlation of awareness with CC effect

# summarise CC effect
sub_CC <- 
  test_diff %>% 
  group_by(subj) %>% 
  summarise(CC_RT = mean(CC_RT))

#summarise awareness
sub_Aware <- 
  data_gen_analysis %>% 
  group_by(subj,TT) %>% 
  summarise(acc = mean(acc)) %>% 
  pivot_wider(names_from = TT, values_from = acc) %>% 
  mutate(CC_aware = Random - Repeated) %>% 
  select(subj, CC_aware) %>% 
  mutate(subj = as.factor(subj))

# join and look at correlation
c_data <- 
  full_join(sub_CC, sub_Aware, by = "subj")

# test correlation
cor(c_data$CC_RT, c_data$CC_aware)
correlationBF(c_data$CC_RT, c_data$CC_aware)

awareness_cor_result <- cor.test(~ CC_RT + CC_aware, data = c_data, method = "pearson")

```


```{r inlcude = TRUE}
# conduct ANOVA
anova_awareness <- 
  data_gen_analysis %>% 
  aov_car(acc ~ Error(subj/(TT*targ_depth)), data = .)

# convert variables to factors for BF analysis
data_gen_analysis <- 
  data_gen_analysis %>% 
  mutate(subj = as.factor(subj),
         TT = as.factor(TT),
         targ_depth = as.factor(targ_depth))

# Bayesian analysis on the awareness data
BFanova_awareness <- 
  anovaBF(formula = acc ~ TT*targ_depth + subj,
          whichRandom = "subj",
          data = data_gen_analysis)


```

```{r}

# create new variables with experiment specific names
CCVR04_demog_summary = demog_summary
CCVR04_basic_resp_stats = basic_resp_stats
CCVR04_perc_removed = perc_removed
CCVR04_overall_RT_Means = overall_RT_Means
CCVR04_RT_avgs_figure = RT_avgs_figure
CCVR04_test_figure = test_figure
CCVR04_data_RT_ANOVA = data_RT_ANOVA
CCVR04_data_RT_test_diff = test_diff
CCVR04_RT_test_diff_figure = test_diff_figure
CCVR04_data_ego_RT_ANOVA = data_ego_RT_ANOVA
CCVR04_test_ego_figure = test_ego_figure
CCVR04_data_awareness = data_gen_descriptives
CCVR04_anova_awareness = anova_awareness
CCVR04_BFanova_awareness = BFanova_awareness
CCVR04_awareness_cor = awareness_cor_result

#save these objects for manuscript
save(CCVR04_demog_summary, 
     CCVR04_basic_resp_stats,
     CCVR04_perc_removed,
     CCVR04_overall_RT_Means,
     CCVR04_RT_avgs_figure,
     CCVR04_test_figure,
     CCVR04_data_RT_ANOVA,
     CCVR04_data_RT_test_diff,
     CCVR04_RT_test_diff_figure,
     CCVR04_data_ego_RT_ANOVA,
     CCVR04_test_ego_figure,
     CCVR04_data_awareness,
     CCVR04_anova_awareness,
     CCVR04_BFanova_awareness,
     CCVR04_awareness_cor,
     file = "../../CCVR_ms_1/CCVR04_export.Rdata")

```

